---
title: "sample_markdown"
author: "Brendan"
date: "5 August 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Playing around with R Markdown
# Header sizes depend on the number of hashes you use

***

Guide [here](https://rmarkdown.rstudio.com/authoring_basics.html)

So apparently, you can write \LaTeX\ in markdown. As well as doing a whole heap of pretty cool stuff like saying that $\frac{3}{9} = \frac{1}{3}$, or that $\sqrt{16} = 4$, or just typing $\alpha, \beta, \gamma$ for lolz.

Or
$x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$
$$x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}$$
---

Nice.

You can also put in R code using three back-ticks, and 'echo = FALSE' if you don't want the code to be printed.
```{r, echo=FALSE}
head(mtcars)
```
But if you set echo = TRUE, it will print:
```{r, echo=FALSE}
head(mtcars)
```

Here are some lists:
* blah
* blah
* blah
  + blah (plus for indented parts)
* 1 underscore or asterisk for *italic*
* 2 "         "         " for __bold__
* Superscript uses the hat, ^as per usual^
* ~~Strikethrough~~ uses the tilde
  
Numeric lists:
1. Hi
2. Hi again
  + Hi again
    + And again?
      + And again?

I will cite myself [^1].

## Inline R code
For when you want programmatic inputs. Wrap the inline code in a single back-tick.
I have `5+10` sheep.

## Images
I'll have to have some in this directory (but i should add their formats to the gitignore as they're large) if I want to upload.

```
More verbatim code. Not necessarily in R.
print("Hello world")
```

## Links!
A [linked phrase][id].
Alternatively,
A [linked phrase](http://www.google.com)

## A Plot
```{r qplot, fig.width = 4, fig.height=3, message=FALSE}
library(ggplot2)
summary(cars)
qplot(speed,dist,data=cars) + geom_smooth()
```

<!-- Here we'll put links -->
[id]: http://www.google.com "Hmm"

<!-- Here we'll put references/biblio -->
[^1] Me.
