---
title: "WaveQTL with HMT - Simulation 1 - Generate Gamma, Phi, Effect Size"
author: "Brendan Law"
date: "11/06/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data_path <- "~/Cpp/WaveQTL_HMT/test/dsQTL/output/" # change this when you port it all over to the Masters Git repo
dataset <- "tree_tie_noQT"
geno_select <- 11 # the one used in the demo

# Helper function to get parent indices of a given index
# From "../code/WaveQTL/waveqtl_hmt_test_calc_sumlog.R"
get_parent_indices <- function(indx, tree_root_indx = 1){
  return_indices <- indx
  return_indices <- (return_indices - tree_root_indx + 1) %/% 2

  # root of tree doesn't return any index
  return_indices[which(indx == tree_root_indx)] <- NA_integer_
  return(return_indices)
}
```

## Background

The idea behind this notebook is to generate a sequence of $\gamma_{s,l}$'s for each scale, $s$ and location, $l$. This isn't as straightforward as in the WaveQTL (no HMT) case as the $\gamma_{s,l}$'s are no longer independent of each other (as they are related by the tree prior structure), hence:

$$P(\gamma_{s,l} = m \mid \hat{\pi}, y_{s,l}, g)$$
is no longer a straightforward quantity to calculate, as the posterior distribution of $\gamma_{s,l}$ now depends on all data, $\mathbf{y} = (y_{1,1},\dots,y_{S,L^S})$. Ie. we now have that:

$$P(\gamma_{s,l} = m \mid \hat{\pi}, \mathbf{y}, g)$$

Hence, the approach we take is to simulate instances of the entire gamma vector, $\boldsymbol{\gamma} = (\gamma_{1,1},\dots,\gamma_{S,L^S}$.

We know that:

$$
\gamma_{1,1} =
\begin{cases}
  1 & \text{wp} & P(\gamma_{1,1} = 1 \mid \hat{\pi}, \mathbf{y}, g) := A_{1,1}(1)\\
  0 & \text{wp} & 1 - A_{1,1}(1)
\end{cases}
$$

We use this to simulate a value for $\gamma_{1,1}$, and then let this flow on sequentially to generate the remaining $\gamma$'s. For scale-location $(2,1)$, for example:

$$
\gamma_{2,1} =
\begin{cases}
  1 & \text{wp} & P(\gamma_{2,1} = 1 \mid \gamma_{1,1} = m, \hat{\pi}, \mathbf{y}, g) := B_{(2,1),(1,1)}(1,m)\\
  0 & \text{wp} & 1 - B_{(2,1),(1,1)}(1,m)
\end{cases}
$$
which depends on $m$, the simulated state of $\gamma_{1,1}$. In general:

$$
\gamma_{s,l} =
\begin{cases}
  1 & \text{wp} & P(\gamma_{s,l} = m \mid \gamma_{p(sl)} = n, \hat{\pi}, \mathbf{y}, g) := B_{sl,p(sl)}(m,n)\\
  0 & \text{wp} & 1 - B_{sl,p(sl)}(m,n)
\end{cases}
$$

Luckily, both quantites $A$ and $B$ are output from our model. ($B$ can be calculated as our model outputs as joint and marginal distributions, so by dividing them we get the conditional quantity, $B$):
$$
P(\gamma_{s,l} = m \mid \gamma_{p(sl)} = n, \hat{\pi}, \mathbf{y}, g) = \frac{P(\gamma_{s,l} = m, \gamma_{p(sl)} = n \mid \hat{\pi}, \mathbf{y}, g)}{P(\gamma_{p(sl)} = n \mid \hat{\pi}, \mathbf{y}, g)}
$$

Here's an example working based on tree-level tying:

```{r load_data, echo = F}
a_1 <- as.matrix(read.table(paste0(data_path,dataset,".fph.pp.txt")))[geno_select,]
b_11 <- as.matrix(read.table(paste0(data_path,dataset,".fph.pp_joint_11.txt")))[geno_select,]
b_10 <- as.matrix(read.table(paste0(data_path,dataset,".fph.pp_joint_10.txt")))[geno_select,]
# dim(a_1);dim(b_11);dim(b_10);
# Just take the 1023 numeric values (excl first one as it's the scaling coefficient), from cols 3:1025. Also, we exp() our values as our software returned everything in logs.
# Keep in mind that the first value of b_11, b_10 is just a placeholder - as the top element of the tree has no parent.
a_1 <- exp(as.numeric(a_1[3:1025]))
b_11 <- exp(as.numeric(b_11[3:1025]))
b_10 <- exp(as.numeric(b_10[3:1025]))
```

Now, let's simulate a sequence of $\gamma$'s:
```{r sim_gamma_1}
gamma_seq <- numeric()
set.seed(10)
rand_seq <- runif(1023)
gamma_seq[1] <- ifelse(rand_seq[1] < a_1[1], 1, 0)

for(i in 2:1023){
  indx <- i
  parent_indx <- get_parent_indices(indx)
  parent_gamma <- gamma_seq[parent_indx]
  
  if(parent_gamma == 1){
    numerator <- b_11[indx]
    denominator <- a_1[parent_indx]
  }else if(parent_gamma == 0){
    numerator <- b_10[indx]
    denominator <- 1 - a_1[parent_indx]
  }
  
  post_prob <- numerator/denominator
  
  gamma_seq[i] <- ifelse(rand_seq[i] < post_prob, 1, 0)
  
}
```

```{r plot_attempt1, eval = F}
### RUN INTERACTIVELY!

# Interleaves vectors with two bookend 0s, and a 0 between each current element
# (for spacing of elements like a tree in a plot)
vector_centeriser <- function(vect){
  in_between_zeros <- length(vect) - 1
  
  res_vect <- c(0,vect[1],0)
    for(i in 1:in_between_zeros){
      res_vect <- c(res_vect,vect[i + 1],0)
    }  
  return(res_vect)
}

# Kind of a start to visualising tree coefficients
par(mfrow = c(10,1)
    ,mar = c(2,2,2,2))
plot(vector_centeriser(gamma_seq[1]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[2:3]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[4:7]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[8:15]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[16:31]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[32:63]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[64:127]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[128:255]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[256:511]),type = "h",ylab = "",axes = F)#; axis(1)
plot(vector_centeriser(gamma_seq[512:1023]),type = "h",ylab = "",axes = F)#; axis(1)
```

Now, we have our sequence of $\gamma$'s, we can then simulate values for each $\beta$. We have:
$$ 
\beta_{s,l} =
\begin{cases}
  \text{sample from 3-param t-dist} & \text{ if } \gamma_{s,l} = 1 \\
  0 & \text{ if } \gamma_{s,l} = 0
\end{cases}
$$
For some more details about simulating from a 3-parameter t-distribution, see the short page on that here. (INSERT LINK)

From our HMT output, we are given:
$$
\text{mean1}_{s,l} = a_{s,l} \\
\text{var1}_{s,l} = \frac{b_{s,l}\nu_{s,l}}{\nu_{s,l} - 2} \\
\nu_{s,l} = N = 70, \forall s,l \\
\therefore b_{s,l} = \frac{\text{var1}_{s,l} * (N - 2)}{N}
$$
```{r sim_beta_1}
# Load mean, var outputs from HMT
mean1 <- as.matrix(read.table(paste0(data_path,dataset,".fph.mean1.txt")))[geno_select,]
var1 <- as.matrix(read.table(paste0(data_path,dataset,".fph.var1.txt")))[geno_select,]
# Just take the 1023 numeric values (excl first one as it's the scaling coefficient), from cols 3:1025.
mean1 <- as.numeric(mean1[3:1025])
var1 <- as.numeric(var1[3:1025])

## back out a, b (from t-dist) parameters:
t_nu <- 70
t_a <- mean1
t_b <- var1*(t_nu-2)/t_nu
num_pheno <- length(mean1)

t_sample <- stats::rt(n = num_pheno, df = t_nu)
t_sample_3p <- t_a+(sqrt(t_b)*t_sample)

## Simulate beta, being either 3-param t-dist, or 0
beta_seq <- rep(0,num_pheno)
gamma_1_indx <- which(gamma_seq == 1) 
beta_seq[gamma_1_indx] <- t_sample_3p[gamma_1_indx]
```

Now we'll use an inverse Haar wavelet transform to transform the $\beta$'s back into data space values ($\alpha$'s), and then calculate some sample statistics (mean, variance) from these.
```{r inverse_wlet}
Wmat_1024 = read.table("~/Cpp/WaveQTL_HMT/data/DWT/Wmat_1024",as.is = TRUE)
# W2mat_1024 = Wmat_1024*Wmat_1024

# currently only using rows 2->1024, as not sure what we're doing with scaling coefficient yet.
# Still columns 1->1024, to generate effect sizes at each of the 1024 bases.

### Do we need to take '-ve' or not?
# beta_dataS = as.vector(-matrix(data=beta_seq, nr = 1, nc = 1024)%*%as.matrix(Wmat_1024))
beta_dataS = as.vector(matrix(data=beta_seq, nr = 1, nc = 1023)%*%as.matrix(Wmat_1024)[2:1024,])
plot(beta_dataS, main = "Simulation 1", type = "l")
```
The above is an example of a data space vector. Now we replicate over a number of samples.

```{r mean_var_sampling}
num_samples <- 1000
set.seed(10)
beta_data_samples <- matrix(nrow = num_samples,ncol = 1024)

for(j in 1:num_samples){
  # Generate gamma
  gamma_seq <- numeric()
  
  rand_seq <- runif(1023)
  gamma_seq[1] <- ifelse(rand_seq[1] < a_1[1], 1, 0)
  
  for(i in 2:1023){
    indx <- i
    parent_indx <- get_parent_indices(indx)
    parent_gamma <- gamma_seq[parent_indx]
    
    if(parent_gamma == 1){
      numerator <- b_11[indx]
      denominator <- a_1[parent_indx]
    }else if(parent_gamma == 0){
      numerator <- b_10[indx]
      denominator <- 1 - a_1[parent_indx]
    }
    
    post_prob <- numerator/denominator
    
    gamma_seq[i] <- ifelse(rand_seq[i] < post_prob, 1, 0)
  }
  
  # Generate beta
  t_sample <- stats::rt(n = num_pheno, df = t_nu)
  t_sample_3p <- t_a+(sqrt(t_b)*t_sample)
  
  ## Simulate beta, being either 3-param t-dist, or 0
  beta_seq <- rep(0,num_pheno)
  gamma_1_indx <- which(gamma_seq == 1) 
  beta_seq[gamma_1_indx] <- t_sample_3p[gamma_1_indx]
  
  # Transform into data space
  beta_data_samples[j,] = as.vector(matrix(data=beta_seq, nr = 1, nc = 1023)%*%as.matrix(Wmat_1024)[2:1024,])
}

sample_mean <- apply(beta_data_samples,MARGIN = 2,mean)
sample_sd <- apply(beta_data_samples,MARGIN = 2,sd)

```
Plots:

```{r sample_plots}
plot(sample_mean,main="sample mean",type="l")
abline(h = 0, col = "red")
plot(sample_sd,main="sample sd",type="l")
abline(h = 0, col = "red")
```

Plots in WaveQTL style:
```{r sample_plots_waveqtl}
ymin_beta = min(sample_mean - 3*sample_sd) - abs(min(sample_mean - 3*sample_sd))*0.0000000001
ymax_beta = max(sample_mean + 3*sample_sd) + abs(max(sample_mean + 3*sample_sd))*0.0000000001

beta_l = sample_mean - 3*sample_sd
beta_r = sample_mean + 3*sample_sd

wh_l = which(beta_l > 0)
wh_r = which(beta_r < 0)
high_wh = sort(unique(union(wh_l, wh_r)))

xval = 1:1024
col_posi = xval[high_wh]

# pdf("../test/dsQTL/effectSize.pdf", width = 8, height=3)
par(mar = c(2,4,4,2))
plot(1,1,type="n", xlab = "position", ylab = "Effect size",ylim=c(ymin_beta, ymax_beta),xlim=c(1, 1024),main ="Posterior mean +/-3 posterior standard deviation", axes=FALSE)
axis(2)
if(length(col_posi) > 0){
for(j in 1:length(col_posi)){
	polygon(c(col_posi[j]-0.5, col_posi[j]-0.5, col_posi[j]+0.5, col_posi[j]+0.5), c(ymin_beta-2, ymax_beta+2, ymax_beta+2, ymin_beta-2), col ="pink", border = NA)
}
}

abline(h = 0, col = "red")
points(xval, sample_mean, col = "blue", type="l")
points(xval, beta_l, col = "skyblue", type="l")
points(xval, beta_r, col = "skyblue", type="l")
box()
```
