---
title: "tdist_3_param"
author: "Brendan Law"
date: "16 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


## The three-parameter (generalised) student's t-distribution 

The student's t-distribution (one parameter, degrees of freedom) can be generalised to a three-parameter location-scale family:

$$ X = \mu + \sigma T$$ or

$$ T = \frac{X - \mu}{\sigma}$$
where $\mu$ is a location parameter, $\sigma$ is a scale parameter, and T is a random variable with the classic student's t-distribution.

For more details, you can view: <https://en.wikipedia.org/wiki/Student%27s_t-distribution#Generalized_Student's_t-distribution>.

Hence, sampling from the 3-parameter t-distribution should be fairly straightforward - use R's inbuilt functions to sample from the t-distribution, then scale by the other two parameters.

### Sampling from a classic t-distribution
For example, take the following spec:
$$ \nu = 10, \mu = 5, \sigma = 2 $$
```{r classic_t}
nu = 10
mu = 5
sigma = 2
t_sample <- stats::rt(n = 1000, df = nu)
t_sample_3p <- mu+(sigma*t_sample)
hist(t_sample)
hist(t_sample_3p)
ggplot() +
  geom_histogram(data = data.frame(t_sample),aes(x = t_sample, alpha = 0.5, fill = "classic")) +
  geom_histogram(data = data.frame(t_sample_3p),aes(x = t_sample_3p, alpha = 0.5, fill = "3param"))
```
