---
title: "waveqtlHMT_effect_size"
author: "Brendan Law"
date: "17 May 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Estimating data-space effect size properties in WaveQTL with HMT

The idea here is to simulate the key statistical properties (mean, variance), of the effect size in the data space. We know that the posterior distribution of $\beta_{s,l}$ follows a mixture distribution of a 3-parameter t-distribution, and a point mass at zero, with mixing probability $\phi$. See the paper/derivations for more details.

Unfortunately, due to the dependency between the mixing probabilities imposed by the HMT priors, we no longer have independent posterior distributions of $\beta_{s,l}$. Therefore the closed form of the means and variances in the data space aren't available in closed form, like they were in the non-HMT case. We have the following simulation procedure:

1. Generate $\beta_{s,l}$ posterior simulations:
$$ cases stuff$$
2. Transform simluations from posterior into data space simulations using inverse Haar wavelet matrix
3. Repeat steps 1 and 2 (say, 1000 times), until we can calculate a sampling mean and variance for each data space parameter, $\alpha_b$, for $b = 1,\dots,B$.

```{r cars}
summary(cars)
```
